Estimating the number of clusters using Cross Validation

### 2015-06-18 cvclust.tex
more progress on proof

\newcommand{\CV}{\operatorname{CV}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\dataX}{\mathfrak{X}}
\newcommand{\Xtrain}{X_{\text{train}}}
\newcommand{\Ytrain}{Y_{\text{train}}}
\newcommand{\Xtest}{X_{\text{test}}}
\newcommand{\Ytest}{Y_{\text{test}}}
\newcommand{\hGX}{\hat G^{X}}
\newcommand{\hGY}{\hat G^{Y}}
\newcommand{\bpi}{\bar \pi}
\newcommand{\bmuX}{\bar \mu^{X}}
\newcommand{\bmuY}{\bar \mu^{Y}}
\newcommand{\sA}{\mathcal{A}}
\newcommand{\sE}{\mathcal{E}}
\newcommand{\sX}{\mathcal{X}}

@@

>>>>

\newcommand{\CV}{\operatorname{CV}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\dataX}{\mathfrak{X}}
\newcommand{\Xtrain}{X_{\text{train}}}
\newcommand{\Ytrain}{Y_{\text{train}}}
\newcommand{\Xtest}{X_{\text{test}}}
\newcommand{\Ytest}{Y_{\text{test}}}
\newcommand{\hGX}{\hat G^{X}}
\newcommand{\hGY}{\hat G^{Y}}
\newcommand{\bpi}{\bar \pi}
\newcommand{\bmuX}{\bar \mu^{X}}
\newcommand{\bmuY}{\bar \mu^{Y}}
\newcommand{\sA}{\mathcal{A}}
\newcommand{\sC}{\mathcal{C}}
\newcommand{\sE}{\mathcal{E}}
\newcommand{\sN}{\mathcal{N}}
\newcommand{\sX}{\mathcal{X}}

@@

\begin{lemma}

If $\int \|x\|^2 P(dx) < \infty$ and the support of $P$ contains at least $k$
points, then $m_1(P) > m_2(P) > \dotsb > m_k(P)$.  Further, there exists a
finite $M$ such that $W(\cdot, P)$ attains its infimum over $\sA_k$ and each
minimizer $\bar A$ is contained in the set
\(
  \sE_k = \{ A \in \sA_k : A \subset B(5M) \},
\)
where $B(5M)$ denotes the closed ball of radius~$5M$.
\end{lemma}
\begin{proof}
This argument is adapted from the proof in Section~3 of Pollard's paper.  We
first find an $M$ large enough to ensure that every minimizer of $W(\cdot, P)$
must have at least one point in $B(M)$.  Next, we impose additional
constraints on $M$ to ensure that all minimizers of $W(\cdot, P)$ are
contained in $B(5M)$.

Pick~$r$ so that the ball~$K$ of radius~$r$ centered at the origin has
positive $P$~measure.  Take $M$ to be large enough such that
\(
  (M - r)^2 P(K) > \int \|x\|^2 P(dx).
\)
Let $\sC_1$ denote the class of sets $A$ with at most $k$ points and at least
one point in in $B(M)$.  Let $A_0$ be the singleton set containing the origin.
If no point of $A$ is in $B(M)$, then
\(
  W(A, P) \geq (M - r)^2 P(K) > W(A_0, P).
\)
Put concisely, if $A \in \sA_k \setminus \sC_1$, then
\(
  W(A, P) > W(A_0, P) \geq m_k(P).
\)

Next, we proceed by induction on $k$.  If $k = 1$, then there is a unique
minimizer of $W(\cdot, P)$ over $\sA_k$, the set containing the mean $\int x
P(dx)$.  For $k > 1$, take as given that $m_{k-1}(P) = W(\bar A_{k-1}, P)$ for
some set $\bar A_{k-1} \in \sE_{k-1}$.  We will show that there is an $M$
large enough so that $W(\cdot, P)$ attains its infimum over $\sA_k$ and all
minimizers lie in $\sE_k$.

We first show that if $M$ is large enough, then there exists a set $A \in \sE_k$
and a positive $\varepsilon$ with $\varepsilon + W(A,P) < m_{k-1}(P)$.
Since the support of $P$ contains least $k$ points, there exists an $x_0$ not
in $\bar A_{k-1}$ that lies in the support of $P$.  Ensure that $M$ is large
enough so that $x_0 \in B(5M)$.  Take $A \in \sE_k$ to be the set of $k$
points gotten by adjoining $x_0$ to $\bar A_{k-1}$.  Let
\(
  \delta = \min_{a \in \bar A_{k-1}} \| x_0 - a \|
\)
and set $R = \{ x \in \mathbb{R}^d : \|x - x_0\| < \delta/3 \}$; since $x_0$
is in the support of $P$, the neighborhood $R$ has positive $P$~measure.
Further,
\[
  \int_{R} \min_{a \in \bar A_{k-1}} \|x - a\|^2 P(dx)
  \geq
  \int_{R} \min_{a \in \bar A_{k-1}} (\|x_0 - a\| - \delta/3)^2 P(dx)
  = (4/9) \delta^2 P(R),
\]
and
\[
  \int_{R} \min_{a \in A} \|x - a\|^2 P(dx)
  =
  \int_{R} \|x - x_0\|^2 P(dx)
  \leq
  (1/9) \delta^2 P(R).
\]
Thus, if
\(
  \varepsilon < (3/9) \delta^2 P(R)
\)
then
\(
  m_{k-1}(P) - W(A, P) = 
  W(\bar A_{k-1}, P) - W(A, P) > \varepsilon.
\)
